---
title: "Covid91 vaccine study Final2023"
author: "Jayline Zelaya"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


## Males

$H_0$ The Covid-921 vaccine does not exhibit a significant difference in effectiveness between male populations.


$H_A$ The Covid-921 vaccine does exhibit a significant difference in effectiveness between male populations.


```{r}
library(dplyr)
Males = filter(FinalData, FinalData$sex=='M')
```


# Methods

Male is categorical variable, Covid-921 vaccine administration is an explanatory variable, Covid-19 prevention effectiveness is the response variable. Therefore, we use cat~cat helper file.



# Results

## Descriptive Results 

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")
```


Of the males who were infected with Covid-19 over 60% were given the placebo and about 30% were given the real Covid921 vaccine. Of the males who were not infected with the Covid-19 virus about half were given the Covid921 vaccine and the other half were given a placebo vaccine.

The frequency of those males infected with Covid-19 is less than 500, among those males who were infected the ones who received a placebo had higher frequency than those who received the vaccine. The frequency of those males who did not get infected with Covid-19 is about 10000, the frequency of males who were not infected both reach over 8000 regardless of whether they took a vaccine or placebo. Although, of those males who were not infected the ones who took a placebo have a higher frequency than the males who were not infected and got the Covid921 vaccine.



### Numerical Descriptive Results

```{r}
tableM <- xtabs(~infected + treatment, data=Males)
rowPerc(tableM)
colPerc(tableM)
```



About 70% of males who received a placebo were infected with Covid-19. Approximately 30% of males that received the real vaccine were infected with Covid-19. 50.48% of males those who received the placebo reported not being infected by the Covid-19 virus, while approximately 50% of males who received the Covid921 vaccine reported feeling Ok and not infected with the Covid19 virus.

In a different clinical trial run 3% of males that received the placebo vaccine were infected with Covid-19, 1% of males who received the actual Covid921 vaccine reported being infected with Covid-19. On the other hand, about 97% of men who received a placebo vaccine were not infected with the Covid-19 virus, and 98.9% of males who received the real Covid921 vaccine reported not being infected by the Covid virus. 


## Inferential Results

```{r}
chisq.test(tableM)
chisqtestGC(tableM)
```



The chisquare test determined p value is less than 0.05 therefore we reject the null hypothesis.





```{r}
#Males$infected=factor(Males$infected,c('none','Covid19'))
fisher.test(Males$treatment,Males$infected)
```

 
P value is .00000000000004631, The odds of the vaccine working for males is about 2.2:1


# Discussion and Conclusion

The original question was does Covid-921 vaccine exhibit a significant difference in effectiveness between male populations. Male is categorical variable, Covid-921 vaccine administration is an explanatory variable, Covid-19 prevention effectiveness is the response variable. Therefore, we use cat~cat helper file. Of the males who were infected with Covid-19 over 60% were given the placebo and about 30% were given the real Covid921 vaccine. Of the males who were not infected with the Covid-19 virus about half were given the Covid921 vaccine and the other half were given a placebo vaccine. About 70% of males who received a placebo were infected with Covid-19. Approximately 30% of males that received the real vaccine were infected with Covid-19. 50.48% of males those who received the placebo reported not being infected by the Covid-19 virus, while approximately 50% of males who received the Covid921 vaccine reported feeling Ok and not infected with the Covid19 virus. The chisquare test determined p value is less than 0.05 therefore we reject the null hypothesis.  The odds of the vaccine working for males is about 2.2:1

## Females

$H_0$ The Covid-921 vaccine does not exhibit a significant difference in effectiveness between female populations.


$H_A$ The Covid-921 vaccine does exhibit a significant difference in effectiveness between female populations.



```{r}
library(dplyr)
Females = filter(FinalData, FinalData$sex=='F')
```


# Methods

Female is categorical variable, Covid-921 vaccine administration is an explanatory variable, Covid-19 prevention effectiveness is the response variable. Therefore, we use cat~cat helper file.



# Results

## Descriptive Results 


### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")
```


About 38% of women who received the Covid921 vaccine were infected with Covid-19, meanwhile a high 60% of women who received a placebo were infected with Covid-19. 50% of women who received the Covid921 vaccine were not infected and 50% of women who received the placebo vaccine did not get infected with the Covid-19 virus.

The frequency of those females who were infected by Covid-19 is less than 1000, Of those females who were infected with Covid-19 the ones given a placebo have a slightly larger frequency than the females who were infected and given the real Covid921 vaccine. The frequency of the females who were not infected with Covid-19 is approximately 10000, the frequency between women who were not infected and given the placebo and women who were not infected and given the Covid921 vaccine is just about even, although the frequency of the placebo receivers who were not infected is marginally higher.


### Numerical Descriptive Results


```{r}
tableF <- xtabs(~infected + treatment, data=Females)
rowPerc(tableF)
colPerc(tableF)
```


Of those females who were infected with the Covid-19 virus 60% received a placebo and 40% received the real Covid 921 vaccine. Approximately, 50% of females who did receive the Covid921 vaccine and 50% of females who received a placebo did not report being infected with the Covid-19 virus.


In a different trial run, among females who were infected with the Covid-19 virus 2% received a placebo vaccine and about 1.4% received an actual Covid921 vaccine. Among females that were not infected by the Covid-19 virus 98% received a placebo vaccine and 98.6% received an actual Covid921 vaccine. 



## Inferential Results

```{r}
chisq.test(tableF)
chisqtestGC(tableF)
```

The chisquared test determined the P value is less than 0.05 therefore we reject the null hypothesis 



```{r}
#Females$infected=factor(Females$infected,c('none','Covid19'))
fisher.test(Females$treatment,Females$infected)
```

P value is equal to  0.0002162, The odds of the vaccine working for females is about 1.50:1


# Discussion and Conclusion

The original question was does Covid-921 vaccine exhibit a significant difference in effectiveness between female populations. Female is categorical variable, Covid-921 vaccine administration is an explanatory variable, Covid-19 prevention effectiveness is the response variable. Therefore, we use cat~cat helper file. About 38% of women who received the Covid921 vaccine were infected with Covid-19, meanwhile a high 60% of women who received a placebo were infected with Covid-19. 50% of women who received the Covid921 vaccine were not infected and 50% of women who received the placebo vaccine did not get infected with the Covid-19 virus. Of those females who were infected with the Covid-19 virus 60% received a placebo and 40% received the real Covid 921 vaccine. Approximately, 50% of females who did receive the Covid921 vaccine and 50% of females who received a placebo did not report being infected with the Covid-19 virus. The chisquared test determined the P value is less than 0.05 therefore we reject the null hypothesis. The odds of the vaccine working for females is about 1.50:1


## LGBTQ

$H_0$ The COVID-921 vaccine exhibits consistent effectiveness across different sexual orientation subgroups.



$H_A$ The COVID-921 vaccine demonstrates varying effectiveness across different sexual orientation subgroups.



```{r}
library(dplyr)
LGBTQ = filter(FinalData, FinalData$LGBTQ=='gay')
head(LGBTQ)
```



# Methods


Sexual orientation LGTQ+ is a categorical variable, Covid-921 vaccine administration is an explanatory variable, Covid-19 prevention effectiveness is a response variable. Therefore, we use cat~cat helperfile.



# Results


## Descriptive Results 


### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")
```



Of those in the subgroup LGBTQ who were infected with the Covid-19 virus less than 20% received a placebo vaccine and over 80% received the Covid921 vaccine. Of those LGBTQ members who were not infected by Covid-19 50% received a placebo vaccine and 50% received a Covid921 vaccine.



The frequency of those from the LGBTQ community infected with Covid-19 is less than 500, those LGBTQ members who were infected and given the real Covid921 vaccine have a higher frequency than those from the LGBTQ who were infected and only given a placebo vaccine. The frequency of those from the LGBTQ community who were not infected is almost 1500 and is just about even between those who were given the real vaccine and the placebo.


### Numerical Descriptive Results

```{r}
tableG <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(tableG)
colPerc(tableG)
```

Of those from the LGBTQ Community that were infected with the Covid-19 virus approximately 15% received a placebo vaccine and approximately 85% received a Covid921 vaccine. Among those members from the LGBTQ community that were not infected by the Covid-19 virus half were given a placebo vaccine and half were given the Covid921 vaccine.


In a different trial run, among those from the LGBTQ community who were infected by the Covid-19 virus less than 1% were given a placebo vaccine and approximately 3.5% were given the Covid921 vaccine. Among those from the LGBTQ community who were not infected by the Covid-19 virus about 99% received a placebo vaccine and about 96.5% received the Covid921 vaccine.



## Inferential Results
```{r}
chisq.test(tableG)
chisqtestGC(tableG)
```



The chisquare test determined p value is less than 0.05 therefore we reject the null hypothesis


```{r}
#LGBTQ$infected=factor(LGBTQ$infected,c('none','Covid19'))
fisher.test(LGBTQ$treatment,LGBTQ$infected)
```

P value is .00000001029, the odds of the vaccine working for those in LGBTQ community is about 0.2:1


# Discussion and Conclusion

The original question was is Covid-921 vaccine effective among those in different sexual orientation subgroups. Sexual orientation LGTQ+ is a categorical variable, Covid-921 vaccine administration is an explanatory variable, Covid-19 prevention effectiveness is a response variable. Therefore, we use cat~cat helperfile. Of those in the subgroup LGBTQ who were infected with the Covid-19 virus less than 20% received a placebo vaccine and over 80% received the Covid921 vaccine. Of those LGBTQ members who were not infected by Covid-19 50% received a placebo vaccine and 50% received a Covid921 vaccine. Of those from the LGBTQ Community that were infected with the Covid-19 virus approximately 15% received a placebo vaccine and approximately 85% received a Covid921 vaccine. Among those members from the LGBTQ community that were not infected by the Covid-19 virus half were given a placebo vaccine and half were given the Covid921 vaccine. The chisquare test determined p value is less than 0.05 therefore we reject the null hypothesis. The odds of the vaccine working for those in LGBTQ community is about 0.2:1

## Druggies


$H_0$ The COVID-921 vaccine does not demonstrate varying effectiveness among individuals with different levels of drug use.



$H_A$  The COVID-921 vaccine demonstrates consistent effectiveness among individuals with different levels of drug use.


```{r}
library(dplyr)
DrugUser = filter(FinalData, FinalData$DrugUser=='yes')
head(DrugUser)
```

# Methods


Drug use is a categorical variable, Covid-921 vaccine administration is an explanatory variable, Covid-19 prevention effectiveness is a response variable. Therefore, we use cat~cat helperfile.



# Results

## Descriptive Results 


### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")
```


Among Drug User's who became infected with the Covid-19 virus less than 20% were given a placebo and over 80% were given the Covid921 vaccine. Among Drug User's who were not infected by Covid-19 a little more than half were given a placebo vaccine and slightly less than half were given the Covid-921 vaccine.


The frequency of Drug Users who were infected with the Covid-19 virus is less than 500, those who were given the actual vaccine and infected have slightly higher frequency than those Drug Users who were infected and only given a placebo. The frequency of Drug Users who were not infected is over 1500 and out of those Drug Users who were not infected the ones given the placebo have a slightly higher frequency.


### Numerical Descriptive Results

```{r}
tableDU <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(tableDU)
colPerc(tableDU)
```



Among Drug Users who were infected with Covid-19 less than 12% were given a placebo vaccine and approximately 88% were given the Covid921 vaccine. Between Drug Users who were not infected with Covid-19 about 51% were given a placebo vaccine and approximately 88% were given the Covid921 vaccine.


In a different trial run between Drug Users who were infected with the Covid-19 virus 0.4% were given a placebo vaccine and about 3% were given the Covid921 vaccine. Among Drug Users who were not infected with Covid-19 99.6% of drug users were given a placebo and about 97% were given the Covid921 vaccine.



## Inferential Results

```{r}
chisq.test(tableDU)
chisqtestGC(tableDU)
```

The chisquare test determined p value is less than 0.05 therefore we reject the null hypothesis.


```{r}
#DrugUser$infected=factor(DrugUser$infected,c('none','Covid19'))
fisher.test(DrugUser$treatment,DrugUser$infected)
```

P value is .000000000232, the odds of the vaccine working for drug users is about 0.1:1


# Discussion and Conclusion

The original question was does the COVID-921 vaccine demonstrate varying effectiveness among individuals with different levels of drug use. Drug use is a categorical variable, Covid-921 vaccine administration is an explanatory variable, Covid-19 prevention effectiveness is a response variable. Therefore, we use cat~cat helperfile. Among Drug User's who became infected with the Covid-19 virus less than 20% were given a placebo and over 80% were given the Covid921 vaccine. Among Drug User's who were not infected by Covid-19 a little more than half were given a placebo vaccine and slightly less than half were given the Covid-921 vaccine. Among Drug Users who were infected with Covid-19 less than 12% were given a placebo vaccine and approximately 88% were given the Covid921 vaccine. Between Drug Users who were not infected with Covid-19 about 51% were given a placebo vaccine and approximately 88% were given the Covid921 vaccine. The chisquare test determined p value is less than 0.05 therefore we reject the null hypothesis. The odds of the vaccine working for drug users is about 0.1:1

# Overall Results




